---
title: "Diamonds Project"
author: "Tilak Muley"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background and data chosen
The data that was chosen was diamonds. This data set comes built in-the 
tidyverse package.The data diamonds has variables: carat, cut, color, clarity, 
depth, table, price, x, y, and z. The variable carat depends on what weight the 
diamond is in this data set the carat ranges from .20-5.01. The variable cut 
identifies what cut quality the diamond is in ranging from ideal, premium, good,
very good and fair. The variable clarity identifies how clear the diamond is. The
variable depth identifies the depth percentage of the diamond which in this data 
set it ranges from 43.0-79.0. The variable price identifies the price of the 
diamonds in U.S. dollars in this data set the prices ranges from $326-$18823. 
The variable x identifies the length in mm in this data set it ranges from 
0mm–10.74mm. The variable y identifies width in mm which in this data set ranges 
from 0mm–58.9mm. The variable z identifies depth in mm which in this data set 
ranges from 0mm–31.8mm.


## Problem definition
The problem I decided to choose to solve is to find out if there is any 
difference or if its similar on price and carat depending on what type of cut
the diamond has (ideal, premium, good, very good and fair) 

## Importing Data and Library
```{r }
library(dplyr) 
library(knitr)
library(ggplot2)
library(data.table)
library(tidyr)
library(tidyverse)

data("diamonds")

```
## EXPLATORY DATA ANALYSIS 
## Summary of Data
Summary:
```{r }
summary(diamonds)
```

## Data Exploration of diamonds data
First 6 rows of the input data frame:
```{r}
head(diamonds)
```
Last 6 rows of the input data frame:
```{r}
tail(diamonds)
```
Dimensions of the data:
```{r}
dim(diamonds)
```
## DATA MANIPULATION
## Data wrangling, munging and cleaning
To Data wrangle and clean up the data I first used the select() command to get 
only the data of the variables I want for my problem. The variables I wanted was 
cut, carat and price. So I created a new data set called diamondWanted. 
```{r }
diamondWanted = select(diamonds, cut,carat,price)
```
Then I arrange the data set by type of cut so all data for fair together, good 
is together, very good together, premium together, and ideal together
```{r}
diamondWanted = diamondWanted %>% arrange(cut)
diamondWanted
```

## Data Exploration of diamondWanted Data 
First 6 rows of the input Data frame with just variables cut, carat, price:
```{r}
head(diamondWanted)
```
Last 6 rows of the input Data frame with just variables cut, carat, price:
```{r}
tail(diamondWanted)
```

## Data Visualization
Scatter plot:
```{r }
ggplot(data = diamondWanted, aes(x = price, y = carat, color = cut)) + 
  geom_point()
```
Box plot of cut and price:
```{r}
ggplot(data = diamondWanted,aes(x = cut, y = price))+geom_boxplot(fill="purple")
```
Box plot of cut and carat:
```{r}
ggplot(data = diamondWanted,aes(x = cut, y = carat))+geom_boxplot(fill="orange")
```

## Conclusion
In conclusion I found that it isn't that much of difference of carat and price 
depending on the type of cut. Also that the main thing is the carat of the 
diamond not the type of cut that affects the price. Overall I found out the cut 
of the diamond doesn't have much of a impact on the carat or the price of the 
diamond. But I did find that ideal cut is much more spread out along the price 
but the carat of any ideal cut diamond does not exist above 3.5. Also learned 
and found the carat medians of each cut is similar. Also price medians of each 
cut is similar to each type of cut. While max and min was fairly close in price 
between all types of cuts big difference was between max of carat between each
type of cut with fair cut having a bigger max and good cut having the smaller 
max. It was also found the smallest carat belongs to a very good cut diamond 
with a price of $367. While biggest carat belongs to a fair cut diamond with a 
price of $18018. While lowest price on a diamond is $326 on a ideal cut .23 
carat and a premium cut .21 carat. While highest price on diamond is $18823 on a 
premium cut 2.29 carat diamond.

